Spin canonical divisors of log stacky curves

### 2015-06-25 stacky_write_up/spin_cring.tex
revisions to end of intro, g0 r<3 case, minor fixes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% document %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\todo{Say that this follows immediately from the three theorems for the special cases.}

Note that in the course of the proof of Theorem
~\ref{thm:g_0_generators_relations} we will not merely obtain
bounds on the degrees of the generators and relations. In fact, if
the genus $g = 0$ or $g = 1$, the proof yields an explicit system
of generators and the initial ideal of relations. While in the
genus $g \geq 2$ case, the proof does not always yield an explicit
system of generators and relations, if the halfcanonical divisor
$\halfcan$ can be written as $\halfcan' + \Lambda$ with $(\halfcan',
0, X)$ a log spin curve, and a system of generators and relations for
$ \halfcan'$ is known, then our proof provides an inductive
procedure for determining the generators and initial ideal of
relations for $\halfcan$. In fact, such a system of generators and
relations are known for $\halfcan'$ in many cases, as detailed in
Neves ~\cite[ Section III.4]{neves:halfcan}.

The remainder of the paper will be primarily devoted to proving
Theorem ~\ref{thm:general_generators_relations}. The idea of the
proof will be to induct on the order and number of elliptic points.
To this end, we first review important background in Section
~\ref{sec:background}. Then, in Section ~\ref{sec:induction}, we
develop various inductive tools. In Sections ~\ref{sec:g_high},
~\ref{sec:g_1} and ~\ref{sec:g_0}, we prove Theorem
~\ref{thm:general_generators_relations} in the cases that the genus
$g \geq 2, g = 1$ and $g = 0$ respectively. Finally, in Section
~\ref{sec:further-questions}, we pose several questions for future
research.

@@

\section{High genus}
\label{sec:g_high}
We now consider the case when the genus is at least 2. In this case, we are able to bound the degrees and relations of the half canonical ring $R_L$. In this section, we do not obtain explicit presentations of $R_L.$ In contrast, in Section ~\ref{sec:g_1} and Section ~\ref{sec:g_0} we will not only obtain bounds on the generators and relations, but also obtain presentations inductively. The tradeoff is that in the cases of genus 0 and 1, we will have to explicitly deal with base cases, making the induction more involved. We begin with the case that $g \geq 2$.

\ssec{Bounds on Generators and Relations in high genus}

@@

\ssec{Main theorem for high genus}
\label{ssec:main_g_high}

@@

\begin{example}
\label{eg:exception- 1-5}
Let $L' = \frac{2}{5}P_1$ be a half canonical divisor on a stacky curve $\sx$. Let $u \in (P_L)_1$ be any nonzero element. We obtain $\di u|_{P_1} = 0,$ since $2P -2Q \sim 0$ and by Riemann--Roch $\dim_k (P_L)_1 = 1.$ Let $y_3 \in (R_L)_3$ be any nonzero element. We obtain $\di y_3|_{P_1} = - P_1,$ by Riemann Roch, since if $y_3$ did not have a pole at $R$, we would obtain $y_3 \in H^0(\sx,3P-3Q) \cong H^0(\sx, P - Q) \cong 0$ as $P \neq Q$. Finally, let $y_5 \in (R_L)_5$ be an element with $y_5|_{P_1} = -P_1$. Then, we claim there is some $\alpha \in k$ so that
$R_L \cong k[u, y_3, y_5]/(y_3^4 - \alpha uy_5^2).$
The proof of this can be gotten by first using Riemann--Roch and counting the pole order at $P_1$ considerations to check the above relation exists, and that the generators and relation determine a ring whose $k$th graded piece has the same dimension of $R_L$. It is analogous to Example ~\ref{ex:base-377}, and is omittied.

@@

We now elaborate on the computation of the results in 
Table ~\ref{table:g-1-exceptional}. We have already explained the case of 
$L' = P - Q + \frac{1}{5}P_1$ above in Example ~\ref{eg:exception- 1-5}. The other cases are similar.

@@

\begin{example}
\label{eg:exception- 1-e}
Let $L = P - Q + \frac{e- 1}{2e}P_1.$ Here, we claim $R_L \cong k[u, y_3, y_5, \ldots, y_e]/I_e,$ where we define $I_5 = y_3^4 - uy_5^2$ and $I_e$ is inductively defied via Theorem ~\ref{thm:ram_order_ind} so that $in(I_e) = in(I_{e-2})k[u, y_3, \ldots, y_e] + \langle y_e z \mid z \neq y_e, y_{e-2} \rangle.$
Furthermore, $\deg u = 2, \deg y_i = i$.
\todo{Note that here there is a quartic generator of $I$ in degree 12, providing exceptions to minimality by quadrics and an exception to the relation bound when $e = 5$.}
\end{example}
\fi

@@

First, we deal with the case $\sx$ has at least 1 stacky point. By Example ~\ref{eg:exception- 1-5}, if $L' = P - Q + \frac{2}{5}P_1,$ then $R_{\halfcan'}$ is generated in degrees $2,3,$ and $5$ with a single relation in degree 12, is admissible, and satisfies the hypotheses of Theorem ~\ref{thm:ram_order_ind}. While $L'$ itself is an exceptional case, as it has a generator in degree $12 > 2 \cdot 5$, after applying Theorem ~\ref{thm:ram_order_ind}, we see that $P - Q + \frac{3}{7}P_2$ does satisfy the constraints of this theorem, because only relations in degree $\leq 14 = 2 \cdot 7$ are added, while the previous relation lies in degree $12$, which is less than $2 \cdot 7 = 14$. Therefore, the Theorem holds for $L' = P - Q + \frac{3}{7}P_2$. Then, applying Theorem ~\ref{thm:ram_order_ind} $\frac{e-7}{2}$ times shows that the Theorem holds for $L = P - Q + \frac{e- 1}{2e}P_1$.

@@

Now suppose $\delta = 0$, so $\halfcan$ is not necessarily effective.
Let $\sigma = (0; e_1, \ldots, e_r; 0)$ be a signature with $r > 0$
and not one of the exceptional cases in Table
~\ref{table:g-0-exceptional}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% document %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

This theorem is proven in three different genus cases, each of
which use similar inductive processes. The high genus case is
proven in Theorem ~\ref{thm:high_generators_relations}. In Theorem
~\ref{thm:g_1_generators_relations} and Theorem
~\ref{thm:g_0_generators_relations}, we also give the respective
genus one and genus zero results after using a similar induction
process while dealing with more complicated base cases and
exceptional cases.

\begin{rem}
The proofs of the main theorems in the low genus cases not only
obtain bounds on the degrees of the generators and relations, but
also yield explicit systems of generators and the initial ideal of
relations. Furthermore, our proof of the high genus case also
provides an inductive procedure for determining the generators and
initial ideal of relations for $\halfcan'$ when the halfcanonical
divisor $\halfcan$ can be written as $\halfcan' + \Lambda$ with
$(\halfcan', 0, X)$ a log spin curve. In fact, such a system of
generators and relations are known for $\halfcan'$ in many cases,
as detailed in Neves ~\cite[ Section III.4]{neves:halfcan}.
\end{rem}

The remainder of the paper will be primarily devoted to proving
Theorem ~\ref{thm:general_generators_relations}. The idea of the
proof will be to induct on the order and number of hyperelliptic-
fixed points. To this end, we first review important background in
Section ~\ref{sec:background}. Then, we develop various inductive
tools in Section ~\ref{sec:induction}. In Sections ~\ref{sec:g_high},
~\ref{sec:g_1} and ~\ref{sec:g_0}, we prove Theorem
~\ref{thm:general_generators_relations} in the high genus, genus
one, and genus zero cases respectively. Finally, in Section
~\ref{sec:further-questions}, we pose several questions for future
research.

@@

\section{High Genus}
\label{sec:g_high}
We now consider the case when the genus is at least 2. In this case, we are able to bound the degrees and relations of the half canonical ring $R_L$. In this section, we do not obtain explicit presentations of $R_L.$ In contrast, in Section ~\ref{sec:g_1} and Section ~\ref{sec:g_0} we will not only obtain bounds on the generators and relations, but also obtain presentations inductively. The tradeoff is that in the cases of genus 0 and 1, we will have to explicitly deal with base cases, making the induction more involved. We begin with the case that $g \geq 2$.

\ssec{Bounds on Generators and Relations in High Genus}

@@

\ssec{Main Theorem for High Genus}
\label{ssec:main_g_high}

@@

\begin{example}
\label{eg:exception-1-5}
Let $L' = \frac{2}{5}P_1$ be a half canonical divisor on a stacky curve $\sx$. Let $u \in (P_L)_1$ be any nonzero element. We obtain $\di u|_{P_1} = 0,$ since $2P -2Q \sim 0$ and by Riemann--Roch $\dim_k (P_L)_1 = 1.$ Let $y_3 \in (R_L)_3$ be any nonzero element. We obtain $\di y_3|_{P_1} = - P_1,$ by Riemann Roch, since if $y_3$ did not have a pole at $R$, we would obtain $y_3 \in H^0(\sx,3P-3Q) \cong H^0(\sx, P - Q) \cong 0$ as $P \neq Q$. Finally, let $y_5 \in (R_L)_5$ be an element with $y_5|_{P_1} = -P_1$. Then, we claim there is some $\alpha \in k$ so that
$R_L \cong k[u, y_3, y_5]/(y_3^4 - \alpha uy_5^2).$
The proof of this can be gotten by first using Riemann--Roch and counting the pole order at $P_1$ considerations to check the above relation exists, and that the generators and relation determine a ring whose $k$th graded piece has the same dimension of $R_L$. It is analogous to Example ~\ref{ex:base-377}, and is omittied.

@@

We now elaborate on the computation of the results in 
Table ~\ref{table:g-1-exceptional}. We have already explained the case of 
$L' = P - Q + \frac{1}{5}P_1$ above in Example ~\ref{eg:exception-1-5}. The other cases are similar.

@@

\begin{example}
\label{eg:exception-1-e}
Let $L = P - Q + \frac{e- 1}{2e}P_1.$ Here, we claim $R_L \cong k[u, y_3, y_5, \ldots, y_e]/I_e,$ where we define $I_5 = y_3^4 - uy_5^2$ and $I_e$ is inductively defied via Theorem ~\ref{thm:ram_order_ind} so that $in(I_e) = in(I_{e-2})k[u, y_3, \ldots, y_e] + \langle y_e z \mid z \neq y_e, y_{e-2} \rangle.$
Furthermore, $\deg u = 2, \deg y_i = i$.
\todo{Note that here there is a quartic generator of $I$ in degree 12, providing exceptions to minimality by quadrics and an exception to the relation bound when $e = 5$.}
\end{example}
\fi

@@

First, we deal with the case $\sx$ has at least 1 stacky point. By Example ~\ref{eg:exception-1-5}, if $L' = P - Q + \frac{2}{5}P_1,$ then $R_{\halfcan'}$ is generated in degrees $2,3,$ and $5$ with a single relation in degree 12, is admissible, and satisfies the hypotheses of Theorem ~\ref{thm:ram_order_ind}. While $L'$ itself is an exceptional case, as it has a generator in degree $12 > 2 \cdot 5$, after applying Theorem ~\ref{thm:ram_order_ind}, we see that $P - Q + \frac{3}{7}P_2$ does satisfy the constraints of this theorem, because only relations in degree $\leq 14 = 2 \cdot 7$ are added, while the previous relation lies in degree $12$, which is less than $2 \cdot 7 = 14$. Therefore, the Theorem holds for $L' = P - Q + \frac{3}{7}P_2$. Then, applying Theorem ~\ref{thm:ram_order_ind} $\frac{e-7}{2}$ times shows that the Theorem holds for $L = P - Q + \frac{e- 1}{2e}P_1$.

@@

Now suppose $\delta = 0$, so $\halfcan$ is not necessarily effective.
Let $\sigma = (0; e_1, \ldots, e_r; 0)$ be a signature with $r > 0$
and not one of the exceptional cases in Table
~\ref{table:g-0-exceptional}. Note that if $r < 3$, then $\deg 
\lfloor d \halfcan \rfloor < 0$ for all $d \geq 0$ so we have the
trivial case where $R(\sx, \Delta, \halfcan) = k$. Thus, we now
only consider when $r \geq 3$.
